Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2016-10-09 fib2.tex
shortened the main text to 4 pages

\author{Simon Burton}
\affiliation{Centre for Engineered Quantum Systems, School of Physics, 
The University of Sydney, Sydney, Australia}
\author{Courtney G.\ Brell}
\affiliation{Institut f\"{u}r Theoretische Physik, Leibniz Universit\"{a}t Hannover, 
Appelstra\ss{}e 2, 30167 Hannover, Germany}
\author{Steven T.\ Flammia}
\affiliation{Centre for Engineered Quantum Systems, School of Physics, 
The University of Sydney, Sydney, Australia}

@@

Topologically ordered quantum systems in two dimensions show tremendous promise for 
long-term storage and processing of quantum information~\cite{Kitaev2003, Dennis2002, Nayak2008}. 
The topological features of such systems are insensitive to local 
perturbations~\cite{Bravyi2010, Bravyi2011a, Michalakis2013}, and they have quasiparticle excitations 
exhibiting anyonic statistics~\cite{Wilczek1990}. These systems can in general be used as 
quantum memories~\cite{Kitaev2003, Dennis2002} or to perform universal topological 
quantum computation~\cite{Freedman2002, Nayak2008}.

@@

In order to efficiently classically simulate an error-correction protocol for 
a topologically ordered quantum memory, it is necessary to simulate 
the physical noise processes, the decoding algorithm, and the physical recovery operations. 
Decoding algorithms are typically designed to run efficiently on a 
classical computer, but there is generally no guarantee that the 
noise and recovery processes should be classically simulable.
Because of this, almost all of the sizable research effort 
on active quantum error correction for topological systems has focused 
on the case of abelian anyons~\cite{Dennis2002, Duclos-Cianci2010, Duclos-Cianci2010a, Wang2010, Wang2010a, Duclos-Cianci2013, Bravyi2011, Bombin2012, Wootton2012, Anwar2014, Watson2014, Hutter2014a, Bravyi2014, Wootton2015, Fowler2015, Andrist2015}.
Systems of abelian anyons are well suited to studying quantum 
error correction because (at the RG fixed point) noise and 
recovery processes can be efficiently simulated numerically, allowing lattice simulations 
of decoding with over 1 million sites~\cite{Duclos-Cianci2010}. 
Standard algorithms are specifically tailored to exploit the abelian nature 
of these particles, particularly that abelian anyons cannot be used for quantum computation.

@@

Concretely, we consider the simulation of quantum error correction in a two-dimensional lattice 
system with Fibonacci anyons, a class of nonabelian anyons that are universal for quantum 
computation~\cite{Freedman2002, Nayak2008}. Fibonacci anyons are experimentally motivated as the 
expected excitations of the $\nu=\frac{12}{5}$ fractional quantum Hall 
states~\cite{Slingerland2001}, and can be realized in several spin 
models~\cite{Levin2005, Bonesteel2012, Kapit2013, Palumbo2014} and composite 
heterostructures~\cite{Mong2014}.
Any of these kinds of physical systems could be used to perform universal topological quantum computation \cggb{citations?}, and can be modelled by our simulations. Natural sources of noise from thermal fluctuations or external perturbations will be suppressed by the energy gap but must still be corrected to allow for scalable computation.

@@

We consider a 
2-dimensional system with
point-like anyonic quasiparticle excitations, which exhibit exotic statistics under exchange.
The worldlines of these exchanges form \emph{braids}
in (2+1)-dimensions
which act unitarily on the Hilbert space of degrees of freedom known as the \emph{fusion space}.
To describe this space we define
the topological observables of the system: these
are total charge measurements within regions bounded by 
closed loops (Wilson loops).
The possible results of such a charge measurement correspond to different anyon particle types in the model.
These dynamics and observables obey algebraic rules
given by a unitary modular tensor category~\cite{Wang2010b}. 
However, from the perspective of
\emph{topological quantum field theory}
anyons are modeled
as punctures in a surface~\cite{Pfeifer2014}, and it is not
the anyons themselves that braid,
but the manifold in which the anyons reside 
that becomes deformed (twisted) by these braid moves.
This is the perspective we adopt (see \Fref{f:braidloop}).

\begin{figure}[t!]
\begin{center}
    \includegraphics[]{pic-braid-loop.pdf}
\caption{
The deformation of observables corresponding to the implementation of a braid process on three anyons is shown. The configurations before and after are connected by world lines shown in red, while the observable Wilson loops are shown in black.
This process relates two different ways of observing (or fusing) the top two charges.
In general, braiding will twist these loops in different ways following
the deformation of the manifold itself.
\label{f:braidloop}
}
\end{center}
\vspace{-10pt}
\end{figure}

The defining difference between abelian and nonabelian
anyon theories is that in an abelian theory particle content alone uniquely determines the outcome of total charge measurements.
In contrast, outcomes for nonabelian charge measurements depend on the history of the particles as well as their type.
That is, the fusion space is one-dimensional for abelian anyons, and generally larger for nonabelian anyons.
In particular, we consider a system supporting (nonabelian) Fibonacci
anyon excitations, denoted by $\tau$.
Two such anyons can have total charge
that is either $\tau$ or $\vac$ (vacuum), or any superposition of these, and
so the fusion space in this case is 2-dimensional.
We can represent basis states for this space using diagrams 
of definite total charge for the Wilson loops, 
and arbitrary states as linear combinations of these diagrams:
\begin{align*}
\includegraphics[]{pic-fibonacci.pdf}
\end{align*}
For $n$ anyons of type $\tau$, the 
dimension of the fusion space grows asymptotically as $\varphi^n$, where $\varphi=\frac{1+\sqrt{5}}{2}$ 
is the golden ratio.

Observables associated to non-intersecting loops commute, and so
a basis for the space can be built from a maximal
set of disjoint, nested loops.
For three anyons, two
possible ways of nesting loops are related
via $F$-moves:
\begin{align}
\label{e:fmove}
\includegraphics[]{pic-fmove.pdf}
\end{align}
Here we have shown particles with charges $a$, $b$, and $c$
as well as total charge $d$.
The lefthand side shows a state where the anyons with charge $a$ and $b$ are
observed to have total charge $i$.
The righthand side involves a superposition 
over total charge $j$ of the anyons with charge $b$ and $c$. 
The green line serves to linearly
order the anyons, and track
deformations in the underlying space, 
i.e.\ the history of braiding processes.
It also denotes the direction along
which $F$-moves occur. A clockwise braiding process, or \emph{half-twist}, is effected by an $R$-move:
\begin{align}
\label{e:rmove}
\includegraphics[]{pic-rmove.pdf}
\end{align}

These diagrams can be composed (glued)
by regarding the outermost observable
as a charge in another diagram.
This operation distributes multiplicatively
over superpositions, and it also respects the $R$ and $F$-moves.
Conversely, \emph{fusion} of anyons corresponds to 
replacing the interior of a loop by a single charge.

@@

In addition to the fusion space there is also a global degeneracy associated with the topology of the 
manifold on which the anyons reside. We consider systems with the topology of a torus, 
which for the Fibonacci anyons gives rise to a 2-fold degeneracy.
This extra degree of freedom can be thought of being associated with the total anyonic charge (or flux) running through the torus itself. Since there are two different possible charges in the Fibonacci anyon model, the global degeneracy is 2.

@@

Typical thermal noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons. Such a model was analyzed in detail for Ising anyons in Ref.~\cite{Brell2013}.
For convenience, in the numerical results presented in this study we will restrict to a simplified noise model consisting of pair-creation events only, though all of our simulation techniques could be applied directly to the more general families of noise models considered in Ref.~\cite{Brell2013} as desired. It was seen in Ref.~\cite{Brell2013} that the pair-creation-only setting was sufficient to capture the qualitative features of an error-correction simulation for the Ising anyons and we have no reason to expect that this would change when considering the Fibonacci anyon model.

Pair-creation events acting within a single tile do not affect the corresponding observable, and so the only pair-creation events we need explicitly consider act between neighboring pairs of tiles. These are then each associated with an edge of the lattice $\Lambda$, chosen 
uniformly at random to model high-temperature, short timescale thermalizing dynamics.
In studies of codes based on abelian anyon models, the error correction threshold or memory lifetime is typically quoted in terms of an iid noise strength per edge. This measure is not appropriate in a nonabelian anyon setting, where noise processes on neighbouring edges do not commute. We thus measure the error correction threshold and memory lifetime in terms of the average number of noise processes per edge. 
It is easy to see that, in those cases where they are both appropriate, these two methods coincide in the limit of low error rates (see Ref.~\cite{Brell2013} for more details). The thresholds we consider in this paper are on the order of $0.1$ noise processes per edge. In models where both noise strength measures are appropriate, such as the toric code under bit-flip errors, the discrepancy between the two values is around $10\%$ at these strengths.

We consider this treatment of anyon dynamics to be a phenomenological model in that
it neglects any microscopic details of the system.
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.
Additionally, while we model a continuum anyon theory, our analysis can be trivially reinterpreted to apply directly to lattice spin models or other systems where space is fundamentally discrete.

@@

Our error-correction algorithm begins by measuring the charge on 
each tile of the lattice, producing a \emph{syndrome} 
of occupied sites. Following this, it joins nearby occupied tiles into clusters, and measures the total charge within each cluster.
Clusters with trivial charge are discarded, and then nearby clusters are merged (agglomeratively~\cite{Hastie2009}).
The merging process iterates at linearly increasing length scales, at each step measuring total charge and discarding trivial clusters, before concluding when 
there is at most a single cluster remaining (see \Fref{f:decode}).
This decoder is based on a hierarchical clustering algorithm~\cite{Hastie2009, Wootton2015b},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}.

Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so must repeatedly physically query the system to measure these charges.
This is also in contrast to abstract quantum codes, where syndrome extraction can typically be performed in one step.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or 
when it performs a homologically nontrivial operation, thereby altering the 
codespace (and failing to perform successful information recovery).
Note that, as was found in Ref.~\cite{Brell2013}, we expect that 
our qualitative results may be reproduced by most alternative families 
of decoders, although it may not be clear in all cases how to extend these decoders to the nonabelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering 
scheme is compatible with the structure in the noise 
processes that allows us to classically simulate them.

@@

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate connected components 
of average size $O(\log(n))$ and variance $O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected components correspond to sets of anyons that could not have interacted at any 
point in their history. 
We are free to neglect the degrees of freedom associated with braiding between components 
because each component has trivial total charge.
This allows us to simulate the braiding processes within each component separately. 
In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over components. 
Since each 
component has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. 
Of course, this is only a probabilistic statement, and so there may be cases where there exist components with size larger than $O(\log(n))$. However, the likelihood of this is suppressed exponentially in the lattice size $L$~\cite{Grimmett1989}.

@@

In the regime where the combined action of noise and error-correction does not percolate
it is reasonable to expect that the simulation is efficient.
However, with strong enough noise the state in the fusion space will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
Despite this, we use heuristics 
that minimize braid moves, which helps to
track the state beyond the
sub-percolated regime.
This enables simulation of error correction in
the regime around the error-correction threshold for linear lattice sizes up to $L=128$.

@@

In order to track the state of the system we maintain 
a set of disjoint directed curves, 
one for each connected component with trivial total charge.
Unlike previous work~\cite{Brell2013}, this creates a dynamically generated basis for the fusion space 
that allows for tensor factorization of disjoint components.
The application of a pair-creation noise process corresponds to the 
addition of an extra curve to the lattice, supporting
two new anyons.
Following the noise processes, our simulation must 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining (arbitrarily) any curves that intersect that tile, 
and then performing $R$- and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
The details of the algorithm and data structures used for 
this procedure will be discussed elsewhere.
Braiding processes that encircle 
a non-trivial loop of our (toric) manifold can also be 
treated in an analogous way, following e.g.~\cite{Pfeifer2012}.

@@

We can guarantee that error-correction will succeed whenever
the action of noise plus decoder 
does not percolate the anyons over the lattice,
but it is possible that percolated events may still result in no error. 
The connection between the percolation threshold and the error-correction threshold 
is not well understood in general~\cite{Hastings2014}, though it is clear 
that our threshold estimate will be a lower bound for 
the true threshold that may be found if all events 
(including those that have percolated) were simulated.
However, naively computing the effect of the simplest possible homologically nontrivial operations consistently yields either a logical error or a complete failure to return to the codespace (see \Aref{s:homnontrivial}). As such it seems reasonable to suggest that neglecting the full simulation of 
these events does not significantly affect our observed threshold value.

@@

We have demonstrated classical simulation of successful error correction in a universal anyon model. 
Though we have chosen several properties of our model and 
simulation in a convenient way for simplicity,
Ref.~\cite{Brell2013} presents good evidence that it is
unlikely these choices will affect our results qualitatively.
In particular, although we have modeled our logical qubit as 
encoded in the global topological degrees of freedom of our 
system, we could have encoded it in the fusion space of several preferred anyons. 
This situation would be appropriate to model error-correction routines for topological quantum computation. 
Additionally, we expect our results to be stable to changes in details of the noise model and decoding algorithm, again following Ref.~\cite{Brell2013}.

@@

There are several interesting avenues for further research. 
Although it is not completely obvious how to do so, 
applying similar methods to more realistic models such as concrete 
microscopic spin models or models with non-topological features would give 
direct insight into practical error rates needed for topological quantum memories in nonabelian systems. 
It would also be interesting to find \emph{classical} spin models 
whose phase diagram encodes the threshold for error correction in these systems~\cite{Dennis2002}.
Finally, an important extension of this work would be to the simulation of \emph{fault-tolerant} error-correction protocols for nonabelian anyon codes.

@@

We thank A.\ Doherty and R.\ Pfeifer for discussions. 
This work was supported by the ARC via EQuS project number CE11001013, by the US Army Research Office grant numbers W911NF-14-1-0098 and W911NF-14-1-0103, ERC grant QFTCMPS and by the cluster of excellence EXC 201 Quantum Engineering and Space-Time Research. STF also acknowledges support from an ARC Future Fellowship FT130101744.

>>>>

\author{Simon Burton}
\affiliation{Centre for Engineered Quantum Systems, School of Physics, 
The University of Sydney, Sydney, Australia}
\author{Courtney G.\ Brell}
\affiliation{Institut f\"{u}r Theoretische Physik, Leibniz Universit\"{a}t Hannover, 
Appelstra\ss{}e 2, 30167 Hannover, Germany}
\affiliation{Perimeter Institute for Theoretical Physics, 
Waterloo, Canada}
\author{Steven T.\ Flammia}
\affiliation{Centre for Engineered Quantum Systems, School of Physics, 
The University of Sydney, Sydney, Australia}

@@

Topologically ordered quantum systems in two dimensions show great promise for 
long-term storage and processing of quantum information~\cite{Kitaev2003, Dennis2002, Nayak2008}. 
The topological features of such systems are insensitive to local 
perturbations~\cite{Bravyi2010, Bravyi2011a, Michalakis2013}, and they have quasiparticle excitations 
exhibiting anyonic statistics~\cite{Wilczek1990}. These systems can be used as 
quantum memories~\cite{Kitaev2003, Dennis2002} or to perform universal topological 
quantum computation~\cite{Freedman2002, Nayak2008}.

@@

In order to efficiently classically simulate an error-correction protocol for 
a topologically ordered quantum memory, it is necessary to simulate 
the physical noise processes, the decoding algorithm, and the physical recovery operations. 
Decoding algorithms are typically designed to run efficiently on a 
classical computer, but there is generally no guarantee that the 
noise and recovery processes should be classically simulable.
Because of this, almost all of the sizable research effort 
on active quantum error correction for topological systems has focused 
on the case of abelian anyons~\cite{Dennis2002, Duclos-Cianci2010, Duclos-Cianci2010a, Wang2010, Wang2010a, Duclos-Cianci2013, Bravyi2011, Bombin2012, Wootton2012, Anwar2014, Watson2014, Hutter2014a, Bravyi2014, Wootton2015, Fowler2015, Andrist2015}, which can be efficiently simulated due to the fact that they cannot be used for quantum computation.
%Systems of abelian anyons are well suited to studying quantum 
%error correction because (at the RG fixed point) noise and 
%recovery processes can be efficiently simulated numerically, allowing lattice simulations 
%of decoding with over 1 million sites~\cite{Duclos-Cianci2010}. 
%Standard algorithms are specifically tailored to exploit the abelian nature 
%of these particles, particularly that abelian anyons cannot be used for quantum computation.

@@

Concretely, we consider quantum error correction in a two-dimensional 
system with Fibonacci anyons, a class of nonabelian anyons that are universal for quantum 
computation~\cite{Freedman2002, Nayak2008}. Fibonacci anyons are experimentally motivated as the 
expected excitations of the $\nu=\frac{12}{5}$ fractional quantum Hall 
states~\cite{Slingerland2001}, and can be realized in several spin 
models~\cite{Levin2005, Bonesteel2012, Kapit2013, Palumbo2014} and composite 
heterostructures~\cite{Mong2014}.
Any of these physical systems could be used to perform universal topological quantum computation, and can be modelled by our simulations. Natural sources of noise from thermal fluctuations or external perturbations will be suppressed by the energy gap but must still be corrected to allow for scalable computation.

@@

We consider a 
2-dimensional system with
point-like anyonic quasiparticle excitations.%, which exhibit exotic statistics under exchange.
The worldlines of these particles form \emph{braids}
in (2+1)-dimensions
which act unitarily on the Hilbert space of degrees of freedom known as the \emph{fusion space}.
This space is associated with
the topological observables of the system: these
are total charge measurements within regions bounded by 
closed loops.% (Wilson loops).
The possible results of such a charge measurement correspond to different anyon particle types in the model.
These dynamics and observables obey algebraic rules
given by a unitary modular tensor category~\cite{Wang2010b}. In contrast to the approach often taken, we consider the topological observables, rather than the anyons themselves, to be affected by braiding processes (see \Fref{f:braidloop}).
%From the perspective of
%\emph{topological quantum field theory},
%anyons are modeled
%as punctures in a surface~\cite{Pfeifer2014}, 
%and the observables, rather than the anyons themselves, are considered to be deformed by braiding processes (see \Fref{f:braidloop}).
%and it is not
%the anyons themselves that braid,
%but the manifold in which the anyons reside 
%that becomes deformed (twisted) by these braid moves.
%This is the perspective we adopt (see \Fref{f:braidloop}).

\begin{figure}[t!]
\begin{center}
    \includegraphics[]{pic-braid-loop.pdf}
\caption{
The deformation of observables corresponding to the implementation of a braid process on three anyons is shown. Configurations before and after are connected by world lines shown in red, while the charge observables are shown in black.
This process relates two different ways of fusing the top two charges.
%In general, braiding will twist these loops in different ways following
%the deformation of the manifold itself.
\label{f:braidloop}
}
\end{center}
\vspace{-10pt}
\end{figure}

The defining difference between abelian and nonabelian
anyon theories is that in an abelian theory particle content alone uniquely determines the outcome of joint charge measurements.
In contrast, outcomes for nonabelian charge measurements depend on the history of the particles as well as their type.
%That is, the fusion space is one-dimensional for abelian anyons, and generally larger for nonabelian anyons.
We consider a system supporting nonabelian Fibonacci
anyon excitations, denoted by $\tau$.
Two such anyons can have total charge
that is either $\tau$ or $\vac$ (vacuum), or any superposition of these, and
so the fusion space in this case is 2-dimensional.
We can represent basis states for this space using diagrams 
of definite total charge for the Wilson loops, 
and arbitrary states as linear combinations of these diagrams:
\begin{align*}
\includegraphics[]{pic-fibonacci.pdf}
\end{align*}
%For $n$ anyons of type $\tau$, the 
%dimension of the fusion space grows asymptotically as $\varphi^n$, where $\varphi=\frac{1+\sqrt{5}}{2}$ 
%is the golden ratio.

Observables associated to non-intersecting loops commute, and so
a basis for the space can be built from a maximal
set of disjoint, nested loops.
For three anyons, two
possible ways of nesting loops are related
via $F$-moves:
\begin{align}
\label{e:fmove}
\includegraphics[]{pic-fmove.pdf}
\end{align}
Here we have shown particles with charges $a$, $b$, and $c$
as well as total charge $d$.
The lefthand side shows a state where the anyons with charge $a$ and $b$ are
observed to have total charge $i$.
The righthand side involves a superposition 
over total charge $j$ of the anyons with charge $b$ and $c$. 
The green line serves to linearly
order the anyons, and track
%deformations in the underlying space, 
%i.e.\ 
the history of braiding processes (see Ref.~\cite{CurveDiagrams} for a more comprehensive discussion).
%It also denotes the direction along which $F$-moves occur. 
A clockwise braiding process, or \emph{half-twist}, is effected by an $R$-move:
\begin{align}
\label{e:rmove}
\includegraphics[]{pic-rmove.pdf}
\end{align}

These diagrams can be composed (glued)
by regarding the outermost observable
as a charge in another diagram.
This operation distributes over superpositions, and it also respects the $R$- and $F$-moves.
Conversely, \emph{fusion} of anyons corresponds to 
replacing the interior of a loop by a single charge.

@@

In addition to the fusion space there is also a global degeneracy associated with the topology of the 
manifold on which the anyons reside. We consider systems on a torus, 
which for the Fibonacci anyons gives rise to a 2-fold degeneracy.
This extra degree of freedom can be thought of being associated with the total anyonic charge (or flux) running through the torus itself. Since there are two different possible charges in the Fibonacci anyon model, the global degeneracy is 2.

@@

Typical thermal noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons. Such a model was analyzed in detail for Ising anyons in Ref.~\cite{Brell2013}.
For convenience, in the numerical results presented in this study we will restrict to considering pair-creation events only, though all of our simulation techniques could be applied directly to the more general noise considered in Ref.~\cite{Brell2013} as desired. It was seen in Ref.~\cite{Brell2013} that the pair-creation-only setting was sufficient to capture the qualitative features of an error-correction simulation for the Ising anyons and we have no reason to expect that this would change when considering the Fibonacci anyon model.

Pair-creation events acting within a single tile do not affect the corresponding observable, and so the only pair-creation events we need explicitly consider act between neighboring pairs of tiles. These are then each associated with an edge of the lattice $\Lambda$, chosen 
uniformly at random to model high-temperature, short timescale thermalizing dynamics.
In studies of codes based on abelian anyon models, the error correction threshold or memory lifetime is typically quoted in terms of an iid noise strength per edge. This measure is not appropriate in a nonabelian anyon setting, where noise processes on neighbouring edges do not commute. We thus measure the error correction threshold and memory lifetime in terms of the average number of noise processes per edge. 
It is easy to see that, in those cases where they are both appropriate, these two methods coincide in the limit of low error rates, and for the typical noise strengths we encounter in this paper, the discrepancy is around $10\%$ (see Ref.~\cite{Brell2013} for more details). 
%The thresholds we consider in this paper are on the order of $0.1$ noise processes per edge. In models where both noise strength measures are appropriate, such as the toric code under bit-flip errors, the discrepancy between the two values is around $10\%$ at these strengths.

We consider this treatment of anyon dynamics to be a phenomenological model in that
it neglects any microscopic details of the system.
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role. Note also that our analysis applies equally well to either a continuum setting, or to discrete lattice models supporting anyonic excitations.
%Additionally, while we model a continuum anyon theory, our analysis can be trivially reinterpreted to apply directly to lattice spin models or other systems where space is fundamentally discrete.

@@

Our error-correction algorithm begins by measuring the charge on 
each tile of the lattice, producing a \emph{syndrome} 
of occupied sites. Following this, it joins nearby occupied tiles into clusters, and measures the total charge within each cluster.
Clusters with trivial charge are discarded, and then nearby clusters are merged (agglomeratively~\cite{Hastie2009}).
The merging process iterates at linearly increasing length scales, at each step measuring total charge and discarding trivial clusters, before concluding when 
there is at most a single cluster remaining (see \Fref{f:decode}).
This decoder is based on a hierarchical clustering algorithm~\cite{Hastie2009, Wootton2015b},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}. 
The error-correction protocol and simulation procedure are discussed in more depth in the supplementary material.

Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so must repeatedly physically query the system to measure these charges.
%This is also in contrast to abstract quantum codes, where syndrome extraction can typically be performed in one step.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or 
when it performs a homologically nontrivial operation, thereby performing a potential logical error.
%altering the 
%codespace (and failing to perform successful recovery).
%Note that, as was found in Ref.~\cite{Brell2013}, we expect that 
%our qualitative results may be reproduced by most alternative families 
%of decoders, although it may not be clear in all cases how to extend these decoders to the nonabelian setting.
%The advantage of using a clustering decoder is 
%its simplicity and flexibility, and the fact that its clustering 
%scheme is compatible with the structure in the noise 
%processes that allows us to classically simulate them.

@@

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate connected components 
of average size $O(\log(L))$ and variance $O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected components correspond to sets of anyons that could not have interacted at any 
point in their history. 
We are free to neglect the degrees of freedom associated with braiding between components 
because each component has trivial total charge.
This allows us to simulate the braiding processes within each component separately. 
In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over components. 
Since each 
component has size only $O(\log(L))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(L))$. 
Of course, this is only a probabilistic statement, and so there may be cases where there exist components with size larger than $O(\log(L))$. However, the likelihood of this is suppressed exponentially in the lattice size $L$~\cite{Grimmett1989}.

@@

In the regime where the combined action of noise and error-correction does not percolate
it is reasonable to expect that the simulation is efficient.
However, with strong enough noise the state in the fusion space will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
%Despite this, we use heuristics 
%that minimize braid moves, which helps to
%track the state beyond the
%sub-percolated regime.
%This enables simulation of error correction in
Nevertheless, we are able to simulate error-correction in
the regime around the error-correction threshold for linear lattice sizes up to $L=128$.

@@

In order to track the state of the system we represent the state of the fusion space with 
a set of disjoint directed curves, 
one for each connected component with trivial total charge.
Unlike previous work~\cite{Brell2013}, this creates a dynamically generated basis for the fusion space 
that allows for tensor factorization of disjoint components.
A pair-creation noise process corresponds to the 
addition of an extra curve to the lattice, supporting
two new anyons.
Following the noise processes, our simulation must 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining the curves that intersect that tile, 
and then performing $R$- and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
This procedure is discussed in more detail in the supplementary material, and the theoretical framework underlying it is expanded in Ref.~\cite{CurveDiagrams}.
%The details of the algorithm and data structures used for 
%this procedure will be discussed elsewhere.
Braiding processes that encircle 
a non-trivial loop of our (toric) manifold can also be 
treated in an analogous way, following e.g.~\cite{Pfeifer2012}.

@@

We can guarantee that error-correction will succeed whenever
the action of noise plus decoder 
does not percolate the anyons over the lattice,
but it is possible that percolated events may still result in no error. 
The connection between the percolation threshold and the error-correction threshold 
is not well understood in general~\cite{Hastings2014}, though it is clear 
that our threshold estimate would be a lower bound for 
the true threshold that may be found if all events 
(including those that have percolated) were simulated. 
In any case, we do not expect this possible discrepancy to be significant in our simulations.
%However, naively computing the effect of the simplest possible homologically nontrivial operations consistently yields either a logical error or a complete failure to return to the codespace (see \Aref{s:homnontrivial}). As such it seems reasonable to suggest that neglecting the full simulation of 
%these events does not significantly affect our observed threshold value.

@@

We have demonstrated classical simulation of successful error correction in a universal anyon model. 
Though we have chosen several properties of our model and 
simulation in a convenient way for simplicity,
Ref.~\cite{Brell2013} presents good evidence that it is
unlikely these choices will affect our results qualitatively.
In particular, although we have considered a logical qubit encoded in the global topological degrees of freedom of our 
system, we could have encoded it in the fusion space of several preferred anyons. 
This situation would be appropriate to model error-correction routines for topological quantum computation. 
Additionally, we expect our results to be stable to changes in details of the noise model and decoding algorithm, again following Ref.~\cite{Brell2013}.

@@

There are several interesting avenues for further research. 
Although it is not completely obvious how to do so, 
applying similar methods to more realistic models such as concrete 
microscopic spin models or models with non-topological features would give 
direct insight into practical error rates needed for nonabelian topological quantum memories.% in nonabelian systems. 
It would also be interesting to find \emph{classical} spin models 
whose phase diagram encodes the threshold for error correction in these systems~\cite{Dennis2002}.
Finally, an important extension of this work would be to the simulation of \emph{fault-tolerant} error-correction protocols for nonabelian anyon codes.

@@

We thank A.\ Doherty and R.\ Pfeifer for discussions. 
This work was supported by the ARC via EQuS project number CE11001013, by the US Army Research Office grant numbers W911NF-14-1-0098 and W911NF-14-1-0103, ERC grant QFTCMPS and by the cluster of excellence EXC 201 Quantum Engineering and Space-Time Research. STF also acknowledges support from an ARC Future Fellowship FT130101744. This research was supported in part by Perimeter Institute for Theoretical Physics. Research at Perimeter Institute is supported by the Government of Canada through Industry Canada and by the Province of Ontario through the Ministry of Research and Innovation.
